\documentclass[letterpaper, 10 pt, conference]{ieeeconf}

% IEEE conference packages
\IEEEoverridecommandlockouts
\overrideIEEEmargins

% Packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{xcolor}

% Theorem environments
\usepackage{amsthm}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}

% Custom commands
\newcommand{\SO}{\mathrm{SO}(3)}
\newcommand{\SE}{\mathrm{SE}(3)}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\DeclareMathOperator*{\argmin}{arg\,min}

% Title information
\title{\LARGE \bf
Bounded-Error Attitude Smoothing on SO(3) via Sequential Convexification and Structured ADMM
}

\author{Author One$^{1}$ and Author Two$^{2}$%
\thanks{$^{1}$Author One is with Department of Electrical Engineering, University Name, City, Country. Email: author1@university.edu}%
\thanks{$^{2}$Author Two is with Department of Robotics, University Name, City, Country. Email: author2@university.edu}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

We study batch attitude smoothing on $\SO$ under bounded measurement uncertainty, where each estimate is constrained by a geodesic tube derived from sensor error limits. The proposed solver combines sequential convexification with trust-region control and a structured ADMM inner loop that exploits block-banded Hessian structure. This yields a practical bounded-error smoothing pipeline that reports tube excess explicitly rather than relying only on soft penalties. Our analysis gives local convergence conditions and an $O(\Delta^2)$ bound on nonlinear mismatch induced by linearization. Experiments on synthetic trajectories and EuRoC MAV sequences demonstrate practical behavior in realistic inertial sensing settings. Relative to generic SOCP baselines, the implementation achieves 5--10$\times$ runtime speedups while keeping tube excess small compared with sensor-derived bounds. Code and reproducible scripts are available at \url{https://github.com/username/SO3-smoothing}.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}

Rotation smoothing is a fundamental problem in robotics and computer vision, where sequences of noisy rotation measurements must be denoised while preserving smooth motion. Applications span camera pose estimation, attitude control, inertial navigation, and 3D reconstruction, where rotation data comes from sensors such as IMUs, visual odometry, or structure-from-motion pipelines.

Traditional approaches to rotation smoothing typically assume Gaussian noise models and minimize energy functions penalizing angular velocity and acceleration. However, many practical scenarios involve bounded-error constraints where sensor specifications, calibration uncertainty, or safety requirements mandate that smoothed estimates remain within known geodesic bounds from the measurements. This set-membership formulation, first introduced in optimal estimation theory for linear systems, has received limited attention for nonlinear manifolds such as $\SO$.

The challenges in $\SO$ attitude smoothing are threefold. First, the nonlinear manifold structure precludes direct application of linear smoothing techniques. Second, tube constraints yield a non-convex constrained problem when posed on the original manifold. Third, practical estimators must balance computational efficiency with explicit bounded-error diagnostics across long trajectories.

\subsection{Contributions}

We present a bounded-noise tube smoothing algorithm for $\SO$ that addresses these challenges. Our \textbf{primary contribution} is a \textbf{reproducible bounded-error attitude smoothing pipeline} that bridges fast manifold solvers and explicit bound-aware estimation.

Specifically, we contribute:

\textbf{1. Bounded-error attitude smoothing formulation.} We formulate smoothing as a set-membership problem on $\SO$ with per-sample geodesic tube constraints, preserving explicit sensor-bound structure instead of replacing it with soft penalties.

\textbf{2. Sequential convexification with mismatch control.} We use local retractions with trust-region management, and provide local convergence conditions together with an explicit bound on nonlinear constraint mismatch caused by linearization.

\textbf{3. Structured ADMM implementation for large batch problems.} We design a custom ADMM solver that exploits block-banded Hessian structure for efficient sparse linear solves, with warm starts and reusable linear algebra components.

\textbf{4. Empirical validation with explicit bound diagnostics.} We report runtime, accuracy, and tube-excess behavior on synthetic data and EuRoC MAV sequences, with ablations and baseline comparisons.

\subsection{Key Results}

Our implementation achieves:
\begin{itemize}
\item \textbf{Small tube excess:} Excess over tube bounds is typically $<10^{-3}$ rad on synthetic tests
\item \textbf{Practical runtime:} 0.5--10 s for windows with $M=100$--$1000$ attitudes
\item \textbf{Speedup over baselines:} 5--10$\times$ faster than generic SOCP solvers
\item \textbf{Real-data behavior:} Evaluated on 5 EuRoC MAV sequences with stable runtime
\end{itemize}

\subsection{Paper Organization}

Section~\ref{sec:related} reviews related work. Section~\ref{sec:methods} presents our formulation and solver. Section~\ref{sec:theory} provides theoretical analysis. Section~\ref{sec:experiments} describes experiments. Section~\ref{sec:results} presents results. Section~\ref{sec:conclusion} concludes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{sec:related}

\subsection{Attitude Estimation on $\SO$}

Attitude estimation on $\SO$ has a long history in guidance, navigation, and tracking. Classical and nonlinear filters remain central in practice, including complementary-filter formulations directly posed on $\SO$~\cite{mahony2008} and modern invariant filtering analyses~\cite{barrau2017}. A broad overview of nonlinear attitude estimation methods is given in~\cite{crassidis2007}. In parallel, batch optimization and smoothing pipelines on manifolds have become standard in robotics and navigation, including manifold preintegration methods for inertial sensing~\cite{forster2015}.

\subsection{Rotation Averaging and Manifold Optimization}

Batch orientation estimation is closely related to rotation averaging, where substantial progress has been made on robust formulations and scalable optimization~\cite{hartley2013,dellaert2020}. These methods provide strong tools for estimating consistent orientation trajectories from relative measurements, but they are typically not formulated as per-sample bounded-error tube constraints tied to sensor-derived limits. Our work is complementary: we focus on bounded tube compliance with smoothness regularization in a direct smoothing formulation.

\subsection{Bounded-Error and Set-Membership Estimation}

Bounded-error (set-membership) estimation provides a deterministic alternative to Gaussian assumptions when uncertainty is more naturally specified by known error bounds~\cite{walter1990,jaulin2001}. This viewpoint is well established in robust estimation and identification, but less common in practical $\SO$ trajectory smoothing workflows that must also scale to long sequences. Our formulation adopts this bounded-error perspective using geodesic tube constraints, and our solver design targets computational efficiency for large batch problems.

\subsection{Positioning of This Work}

Compared with unconstrained or soft-penalty smoothers, we preserve explicit per-sample tube bounds throughout sequential convexification and report tube excess explicitly in experiments. Compared with generic conic optimization approaches, we exploit block-banded structure with a tailored ADMM implementation inspired by structured large-scale optimization practice~\cite{boyd2011,boyd_scp}. The resulting contribution is primarily methodological and computational: a reproducible bounded-error $\SO$ smoothing pipeline with practical runtime gains and bounded linearization-mismatch analysis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
\label{sec:methods}

\subsection{Problem Formulation}

Given a sequence of noisy rotation measurements $R_{\text{meas},i} \in \SO$ for $i = 0, \ldots, M-1$, and per-measurement tube radii $\epsilon_i > 0$, we seek smoothed rotations $R_i \in \SO$ that:

\begin{enumerate}
\item Satisfy tube constraints: $\|\log(R_{\text{meas},i}^T R_i)\|_2 \leq \epsilon_i$
\item Minimize smoothness cost (angular velocity and acceleration)
\end{enumerate}

We parameterize rotations using the exponential map:
\begin{equation}
R_i = \exp(\phi_i^\wedge), \quad \phi_i \in \mathbb{R}^3
\end{equation}

The optimization problem is:
\begin{align}
\min_{\phi \in \mathbb{R}^{3M}} \quad & \frac{1}{2} \phi^T H \phi \label{eq:objective} \\
\text{s.t.} \quad & \|r_i(\phi_i)\|_2 \leq \epsilon_i, \quad i = 0, \ldots, M-1 \nonumber
\end{align}

where $H$ is the smoothness Hessian and $r_i(\phi_i) = \log(R_{\text{meas},i}^T \exp(\phi_i^\wedge))$.

\subsection{Measurement-Bound Construction}

To connect the tube radii to sensor specifications, we model per-step attitude uncertainty from three components: gyro white noise, bias uncertainty, and residual calibration error. For sampling interval $\Delta t_i$, we define
\begin{equation}
\sigma_{\theta,i}^2
=
(\sigma_g \sqrt{\Delta t_i})^2
+
(\sigma_b \Delta t_i)^2
+
\sigma_c^2,
\end{equation}
where $\sigma_g$ is gyro noise density, $\sigma_b$ is bias uncertainty, and $\sigma_c$ is a calibration residual term (all in consistent units). The geodesic tube radius is then set as
\begin{equation}
\epsilon_i
=
\kappa\,\sigma_{\theta,i},
\end{equation}
with confidence factor $\kappa$ (e.g., $\kappa=3$ for conservative operation). This construction provides a reproducible mapping from sensor-level uncertainty assumptions to per-sample tube constraints used by the smoother.

\subsection{Smoothness Hessian}

The Hessian combines first and second-order differences:
\begin{equation}
H = \frac{\lambda}{\tau} (D_1^T D_1 \otimes I_3) + \frac{\mu}{\tau^3} (D_2^T D_2 \otimes I_3)
\end{equation}

where $D_1$ and $D_2$ are first and second-order difference matrices, $\lambda$ and $\mu$ are weights, and $\tau$ is the time step.

\subsection{Sequential Convexification}

Since constraints are nonlinear, we linearize at each iteration. At iterate $\phi^k$, the constraint becomes:
\begin{equation}
\|r_i + J_i \delta_i\|_2 \leq \epsilon_i
\end{equation}

where $r_i = r_i(\phi_i^k)$ and $J_i = J_r^{-1}(r_i) J_r(\phi_i^k)$ is the Jacobian.

We add trust-region constraints:
\begin{equation}
\|\delta_i\|_2 \leq \Delta
\end{equation}

The subproblem is a convex SOCP solved via ADMM.

\subsection{Structured ADMM Solver}

The ADMM formulation splits the problem:
\begin{align}
\min_{\delta, y, w} \quad & \frac{1}{2} \delta^T H \delta + g^T \delta \\
\text{s.t.} \quad & y_j = r_j + J_j \delta_j, \quad \|y_j\| \leq \epsilon_j \nonumber \\
& w_j = \delta_j, \quad \|w_j\| \leq \Delta \nonumber
\end{align}

ADMM iterations:
\begin{enumerate}
\item $\delta$-update: Solve sparse linear system $(H + \rho A^T A) \delta = -g + \rho A^T (z - u)$
\item $y$-update: Project to $L_2$ balls (parallelizable)
\item $w$-update: Project to trust region (parallelizable)
\item Dual update: Standard ADMM
\end{enumerate}

The linear system is sparse and banded, solved efficiently via sparse LU.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEORY SECTION
\input{sections/theory_detailed}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EXPERIMENTS SECTION
\input{sections/experiments_complete}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}

We presented an efficient implementation of bounded-error attitude smoothing on $\SO$. The method enforces tube constraints in convexified subproblems and reports final tube excess explicitly, providing a practical bounded-uncertainty smoothing tool for batch estimation workflows.

\subsection{Summary}

Our method provides:
\begin{itemize}
\item A set-membership smoothing formulation on $\SO$ with geodesic tube constraints
\item Local mismatch analysis linking trust-region size to nonlinear constraint error
\item Structured ADMM runtime gains (5--10$\times$ versus generic SOCP baselines)
\item Reproducible synthetic and EuRoC evaluations with explicit tube-excess metrics
\end{itemize}

\subsection{Limitations}

\begin{itemize}
\item Convergence to strict tolerances can be slow; practical operation favors relaxed stopping criteria
\item Current scaling evidence is strongest for $M \leq 1000$ in the reported experiments
\item The implementation is single-threaded and does not yet include production-grade real-time integration
\end{itemize}

\subsection{Future Work}
\begin{itemize}
\item Sensor-model-driven construction of tube bounds (including bias and calibration uncertainty)
\item Extension to coupled attitude-dynamics smoothing models and SE(3) state estimation
\item Stronger comparisons with established navigation estimators (e.g., invariant filtering/smoothing baselines)
\item Online and multi-rate variants for tracking pipelines with asynchronous sensing
\end{itemize}

\subsection{Reproducibility}

All code, data, and benchmarks are available at \url{https://github.com/username/SO3-smoothing}. The repository includes:
\begin{itemize}
\item Complete implementation (Python)
\item Reproducible benchmark scripts
\item EuRoC data preprocessing tools
\item LaTeX source for this paper
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}

This work was supported by [funding sources]. We thank the EuRoC MAV dataset authors for providing the real-world validation data.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
